\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2021-09-16 Manuscript/Manuscript.tex
wording edit from reproductive to reproductive biology;
and

describing new model with prsence absence. Now I have to redo analysis of supp. mat to match all analyses

\textbf{Plant life strategies are often delimited by vegetative and
physiological traits but little is known about how floral and
reproductive traits drive these strategies, and in turn shape plant
interactions with floral visitors. Here, we compiled 13 floral, 4
reproductive and 3 vegetative traits for 1,506 plant species from 28
plant-pollinator network studies across 18 different countries. We
investigated the associations among these traits, pollinator visitation
and the functional role of plant species within the networks
(interaction frequency, normalized degree and specialization). We found
that 51.8\% of trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size (PC1, 26.72\%). The second axis indicated a trade-off for the level
of pollinator dependency (PC2, 25.08\%). Although the main axes of trait
variation did not fully explain pollinator visitation rates, different
plant life strategies were associated with visitation rates and
pollinator functional groups. Overall, the main traits that determined
plant species' functional roles were height, nectar concentration,
pollen grains per flower, number of ovules, style length, selfing level
and flower width. Our results highlight the need to consider plant
reproductive and floral traits to improve understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.}

@@

There is an astonishing diversity of floral structures and plant
reproductive strategies among flowering
plants\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}},
which have long been of interest to pollination biologists in terms of
their relevance to plant-pollinator interactions. However, most studies
that have explored reproductive (e.g., mating and compatibility systems)
and floral trait (e.g., flower size or nectar provision) variation have
concentrated on the individual or community level and thus, broader
macroecological patterns remain poorly
investigated\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3}--\protect\hyperlink{ref-grossenbacher2017}{7}}.
Indeed, studies depicting species' life history strategies generally
focus on vegetative traits and rarely consider reproductive
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}}. As a
consequence, a unified framework that explores the compromises among
floral traits and their relevance to plant life strategies is currently
lacking\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}}. At the
same time, there is growing interest in the determinants of
plant-pollinator interactions via trait-based
approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{11}} and
trait-matching
analyses\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{12}}.
However, floral traits have been overlooked beyond highly specialised
plant-pollinator
systems\textsuperscript{\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-dellinger2020}{13}}
and the role of plant reproductive biology remains little explored in
plant-pollinator interactions (but see
references\textsuperscript{\protect\hyperlink{ref-tur2013}{14},\protect\hyperlink{ref-devaux2014}{15}}).

With the recent availability of large trait databases, plant ecological
strategies are increasingly being
examined\textsuperscript{\protect\hyperlink{ref-kattge2011}{16},\protect\hyperlink{ref-salguero2015}{17}},
and are facilitating the identification of global patterns and
constraints of plant form and
function\textsuperscript{\protect\hyperlink{ref-salguero2016}{8},\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}}.
However, the main focus has been on vegetative traits such as
leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{20}} or
wood\textsuperscript{\protect\hyperlink{ref-chave2009}{21}} trade-offs
with little or no attention given to reproductive and floral
traits\textsuperscript{\protect\hyperlink{ref-evojtko2020}{22}}, also
critical to plant form and function. For instance, short lived versus
perennial species tend to have low versus high levels of outcrossing,
respectively\textsuperscript{\protect\hyperlink{ref-moeller2017}{6},\protect\hyperlink{ref-barrett2003}{23}}.
Further, outcrossing levels are positively correlated with flower
size\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}}. In
addition, the presence of costly rewards (e.g., pollen or nectar) and
showy flowers or floral displays can only be understood through
consideration of plant species' reliance upon animal pollination
(pollinator dependence) and their role in attracting
pollinators\textsuperscript{\protect\hyperlink{ref-ollerton2011}{25}}.
Hence, exploring plant life strategies with reproductive and floral
trade-offs, in conjunction with their pollinator dependence, is
necessary for a balanced understanding of plant economics.

@@

Here, we explore the potential trade-offs among plant floral and
reproductive traits and how these influence the structure of
plant-pollinator networks. First, we identify the major axes of floral
and reproductive trait variation and trade-offs that determine plant
form and function. Second, we investigate how plant species' position in
trait-space influences their interaction strength with different guilds
of floral visitors. Finally, we investigate how the main axes of trait
variation and individual traits influence plant species roles within
networks.

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\textbf{Fig. 3 \textbar{} Interaction
(yes/no) and visitation rates across the three main axes of trait
variation per floral visitor guild.} Fitted posterior estimates of the
presence/absence of interaction (a, b and c) and number of visits (d, e
and f) made by the different floral visitors guilds in relation to PC1,
PC2 and PC3. PC1 represents the flower number - flower size trade-off,
PC2 represents the pollinator dependence trade-off and PC3, the style
length trade-off.}
\end{figure}

@@

Here, we show that plant species exhibit clear trade-offs in their
floral, reproductive and vegetative traits. These trade-offs are
differentiated on three main axes of trait variation: (i) flower number
- flower size, (ii) pollinator dependence and (iii) style length.
Although these axes of trait variation did not explain overall
visitation rates, we found that plant life strategies were clearly
associated with different floral visitor guilds. Interestingly, pollen
and nectar related traits were better than all other traits for
characterizing plant species functional roles within their interaction
networks.

Over 50\% of plant trait variation was captured by the flower number -
flower size and pollinator dependence trade-offs. Variation in these two
axes of trait variation appears to be associated with the commonly
described `fast-slow continuum' in
plant\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}} and
animal\textsuperscript{\protect\hyperlink{ref-healy2019}{34}}
life-history strategies as indicated by the different floral and
reproductive traits associated with plant height, life form and
lifespan. The `slow' part of the continuum (i.e., tall trees and shrubs)
had plant species with many flowers, few ovules, higher pollinator
dependence, greater occurrence of self-incompatibility and more complex
breeding systems (e.g., monoecious and dioecious species). In contrast,
species that employed the `fast' strategy (i.e., short herbs), had fewer
flowers, more ovules, greater occurrence of self compatibility and lower
pollinator dependence. Further, on the first two axes of trait
variation, we found additional support for the positive association
already described in a previous study between outcrossing rate and
floral
display\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}}.
This trend was further confirmed when looking at the specific
correlation between floral display and autonomous selfing for our
dataset (see Supplementary Fig. S10).

@@

To conclude, we provide the first description of plant floral and
reproductive trade-offs using a large global dataset of plant traits and
interactions with floral visitors. This allowed us to identify the major
reproductive strategies of flowering plants and how plant life
strategies determine their interaction with floral visitor guilds.
Although the percentage of explained variation found with the first two
axes (\textasciitilde{}50\%) is lower than other studies that have
evaluated life histories with vegetative
traits\textsuperscript{\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}},
this percentage is consistent with the largest and most recent study
that has characterized life strategies with both vegetative and
reproductive (i.e., semelparity and iteroparity)
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}}. Future
work needs to integrate these floral and reproductive compromises with
the already described vegetative and physiological
trade-offs\textsuperscript{\protect\hyperlink{ref-diaz2016}{18}} to
create a more comprehensive spectrum of trait
variation\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}}.
Further, the varying level of phylogenetic signal among traits deserves
further attention to understand evolutionary changes on mating and
flower morphology in response to
pollinators\textsuperscript{\protect\hyperlink{ref-gervasi2017}{42},\protect\hyperlink{ref-mackin2021}{43}}.
Finally, a more complete description of the macroecological patterns of
plant-pollinator interactions including unrepresented areas of the
world\textsuperscript{\protect\hyperlink{ref-poisot2021}{44}} and traits
that have been largely ignored (e.g., floral rewards and plant mating
system) are needed to progress in our understanding of the global
patterns in plant-pollinator interactions.

@@

\textbf{Functional traits.} We selected 20 different functional traits
based on their relevance to plant reproduction and data availability
(Table 1). These included twelve quantitative and eight categorical
traits belonging to three broader trait groupings (13 floral, 4
reproductive and 3 vegetative, Supplementary Information). For each
plant species, we undertook an extensive literature and online search
across a wide range of resources (plant databases, online floras, books,
journals and images). From a total of 30,120 cells (20 columns × 1,506
species) we were able to fill 24,341 cells (80.8\% of the dataset, see
Supplementary Fig. S1 for missing values information for each trait).

@@

\caption{\label{tab:unnamed-chunk-5}\textbf{Table 1 | Quantitative and categorical traits used in this study.}}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{>{}cl>{}ccl}
\toprule
\multicolumn{2}{c}{\textbf{Quantitative traits}} & \multicolumn{3}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-5}
\textbf{Type} & \textbf{Traits} & \textbf{Type} & \textbf{Traits} & \textbf{Categories}\\
\midrule
\textbf{Vegetative} & Plant height (m) & \textbf{Vegetative} & Lifepan & \makecell[l]{Short-lived \\ Perennial}\\
\addlinespace
\textbf{Floral} & Flower width (mm) & \textbf{Vegetative} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree}\\
\addlinespace
\textbf{Floral} & Flower length (mm) & \textbf{Floral} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube}\\
\addlinespace
\textbf{Floral} & Inflorescence width (mm) & \textbf{Floral} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic}\\
\addlinespace
\textbf{Floral} & Style length (mm) & \textbf{Floral} & Nectar & \makecell[l]{Presence \\ Absence}\\
\addlinespace
\textbf{Floral} & Ovules per flower & \textbf{Reproductive} & Autonomous selfing & \makecell[l]{None \\ Low \\ Medium \\ High}\\
\addlinespace
\textbf{Floral} & Flowers per plant & \textbf{Reproductive} & Compatibility system & \makecell[l]{Self-incomp. \\ Part. self-comp. \\ Self-comp.}\\
\addlinespace
\textbf{Floral} & Nectar ($\mu$l) & \textbf{Reproductive} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious}\\
\addlinespace
\textbf{Floral} & Nectar (mg) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Nectar concentration ($\%$) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Pollen grains per flower & \textbf{} &  & \\
\addlinespace
\textbf{Reproductive} & Autonomous selfing (fruit set) & \textbf{} &  & \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\textbf{Plant life strategies are often delimited by vegetative and
physiological traits but little is known about how floral and
reproductive biology traits drive these strategies, and in turn shape
plant interactions with floral visitors. Here, we compiled 13 floral, 4
reproductive biology and 3 vegetative traits for 1,506 plant species
from 28 plant-pollinator network studies across 18 different countries.
We investigated the associations among these traits, pollinator
visitation and the functional role of plant species within the networks
(interaction frequency, normalized degree and specialization). We found
that 51.8\% of trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size (PC1, 26.72\%). The second axis indicated a trade-off for the level
of pollinator dependency (PC2, 25.08\%). Although the main axes of trait
variation did not fully explain pollinator visitation rates, different
plant life strategies were associated with visitation rates and
pollinator functional groups. Overall, the main traits that determined
plant species' functional roles were height, nectar concentration,
pollen grains per flower, number of ovules, style length, selfing level
and flower width. Our results highlight the need to consider plant
reproductive and floral traits to improve understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.}

@@

There is an astonishing diversity of floral structures and plant
reproductive strategies among flowering
plants\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}},
which have long been of interest to pollination biologists in terms of
their relevance to plant-pollinator interactions. However, most studies
that have explored plant reproductive biology (e.g., mating and
compatibility systems) and floral trait (e.g., flower size or nectar
provision) variation have concentrated on the individual or community
level and thus, broader macroecological patterns remain poorly
investigated\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3}--\protect\hyperlink{ref-grossenbacher2017}{7}}.
Indeed, studies depicting species' life history strategies generally
focus on vegetative traits and rarely consider reproductive biology
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}}. As a
consequence, a unified framework that explores the compromises among
floral traits and their relevance to plant life strategies is currently
lacking\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}}. At the
same time, there is growing interest in the determinants of
plant-pollinator interactions via trait-based
approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{11}} and
trait-matching
analyses\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{12}}.
However, floral traits have been overlooked beyond highly specialised
plant-pollinator
systems\textsuperscript{\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-dellinger2020}{13}}
and the role of plant reproductive biology remains little explored in
plant-pollinator interactions (but see
references\textsuperscript{\protect\hyperlink{ref-tur2013}{14},\protect\hyperlink{ref-devaux2014}{15}}).

With the recent availability of large trait databases, plant ecological
strategies are increasingly being
examined\textsuperscript{\protect\hyperlink{ref-kattge2011}{16},\protect\hyperlink{ref-salguero2015}{17}},
and are facilitating the identification of global patterns and
constraints of plant form and
function\textsuperscript{\protect\hyperlink{ref-salguero2016}{8},\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}}.
However, the main focus has been on vegetative traits such as
leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{20}} or
wood\textsuperscript{\protect\hyperlink{ref-chave2009}{21}} trade-offs
with little or no attention given to reproductive biology and floral
traits\textsuperscript{\protect\hyperlink{ref-evojtko2020}{22}}, also
critical to plant form and function. For instance, short lived versus
perennial species tend to have low versus high levels of outcrossing,
respectively\textsuperscript{\protect\hyperlink{ref-moeller2017}{6},\protect\hyperlink{ref-barrett2003}{23}}.
Further, outcrossing levels are positively correlated with flower
size\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}}. In
addition, the presence of costly rewards (e.g., pollen or nectar) and
showy flowers or floral displays can only be understood through
consideration of plant species' reliance upon animal pollination
(pollinator dependence) and their role in attracting
pollinators\textsuperscript{\protect\hyperlink{ref-ollerton2011}{25}}.
Hence, exploring plant life strategies with reproductive biology and
floral trade-offs, in conjunction with their pollinator dependence, is
necessary for a balanced understanding of plant economics.

@@

Here, we explore the potential trade-offs among plant floral and
reproductive biology traits and how these influence the structure of
plant-pollinator networks. First, we identify the major axes of trait
variation and trade-offs that determine plant form and function. Second,
we investigate how plant species' position in trait-space influences
their interaction strength with different guilds of floral visitors.
Finally, we investigate how the main axes of trait variation and
individual traits influence plant species roles within networks.

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\textbf{Fig. 3 \textbar{} Interaction
(yes/no) and visitation rates across the three main axes of trait
variation per floral visitor guild.} Fitted posterior estimates of the
presence/absence of interaction (a, b and c) and number of visits (d, e
and f) made by the different floral visitors guilds in relation to PC1,
PC2 and PC3. PC1 represents the flower number - flower size trade-off,
PC2 represents the pollinator dependence trade-off and PC3, the style
length trade-off. For visualization purposes, due to large differences
between the visitation rates of bees and the rest of guilds, the number
of visits was log-transformed (Y-axis of lower panel).}
\end{figure}

@@

Here, we show that plant species exhibit clear trade-offs in their
floral, reproductive biology and vegetative traits. These trade-offs are
differentiated on three main axes of trait variation: (i) flower number
- flower size, (ii) pollinator dependence and (iii) style length.
Although these axes of trait variation did not explain overall
visitation rates, we found that plant life strategies were clearly
associated with different floral visitor guilds. Interestingly, pollen
and nectar related traits were better than all other traits for
characterizing plant species functional roles within their interaction
networks.

Over 50\% of plant trait variation was captured by the flower number -
flower size and pollinator dependence trade-offs. Variation in these two
axes of trait variation appears to be associated with the commonly
described `fast-slow continuum' in
plant\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}} and
animal\textsuperscript{\protect\hyperlink{ref-healy2019}{34}}
life-history strategies as indicated by the different floral and
reproductive biology traits associated with plant height, life form and
lifespan. The `slow' part of the continuum (i.e., tall trees and shrubs)
had plant species with many flowers, few ovules, higher pollinator
dependence, greater occurrence of self-incompatibility and more complex
breeding systems (e.g., monoecious and dioecious species). In contrast,
species that employed the `fast' strategy (i.e., short herbs), had fewer
flowers, more ovules, greater occurrence of self compatibility and lower
pollinator dependence. Further, on the first two axes of trait
variation, we found additional support for the positive association
already described in a previous study between outcrossing rate and
floral
display\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}}.
This trend was further confirmed when looking at the specific
correlation between floral display and autonomous selfing for our
dataset (see Supplementary Fig. S10).

@@

To conclude, we provide the first description of plant floral and
reproductive biology trade-offs using a large global dataset of plant
traits and interactions with floral visitors. This allowed us to
identify the major reproductive strategies of flowering plants and how
plant life strategies determine their interaction with floral visitor
guilds. Although the percentage of explained variation found with the
first two axes (\textasciitilde{}50\%) is lower than other studies that
have evaluated life histories with vegetative
traits\textsuperscript{\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}},
this percentage is consistent with the largest and most recent study
that has characterized life strategies with both vegetative and
reproductive (i.e., semelparity and iteroparity)
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}}. Future
work needs to integrate these floral and reproductive compromises with
the already described vegetative and physiological
trade-offs\textsuperscript{\protect\hyperlink{ref-diaz2016}{18}} to
create a more comprehensive spectrum of trait
variation\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}}.
Further, the varying level of phylogenetic signal among traits deserves
further attention to understand evolutionary changes on mating and
flower morphology in response to
pollinators\textsuperscript{\protect\hyperlink{ref-gervasi2017}{42},\protect\hyperlink{ref-mackin2021}{43}}.
Finally, a more complete description of the macroecological patterns of
plant-pollinator interactions including unrepresented areas of the
world\textsuperscript{\protect\hyperlink{ref-poisot2021}{44}} and traits
that have been largely ignored (e.g., floral rewards and plant mating
system) are needed to progress in our understanding of the global
patterns in plant-pollinator interactions.

@@

\textbf{Functional traits.} We selected 20 different functional traits
based on their relevance to plant reproduction and data availability
(Table 1). These included twelve quantitative and eight categorical
traits belonging to three broader trait groupings (13 floral, 4
reproductive biology and 3 vegetative, Supplementary Information). For
each plant species, we undertook an extensive literature and online
search across a wide range of resources (plant databases, online floras,
books, journals and images). From a total of 30,120 cells (20 columns ×
1,506 species) we were able to fill 24,341 cells (80.8\% of the dataset,
see Supplementary Fig. S1 for missing values information for each
trait).

@@

\caption{\label{tab:unnamed-chunk-5}\textbf{Table 1 | Quantitative and categorical traits used in this study.}}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{>{}cl>{}ccl}
\toprule
\multicolumn{2}{c}{\textbf{Quantitative traits}} & \multicolumn{3}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-5}
\textbf{Type} & \textbf{Traits} & \textbf{Type} & \textbf{Traits} & \textbf{Categories}\\
\midrule
\textbf{Vegetative} & Plant height (m) & \textbf{Vegetative} & Lifepan & \makecell[l]{Short-lived \\ Perennial}\\
\addlinespace
\textbf{Floral} & Flower width (mm) & \textbf{Vegetative} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree}\\
\addlinespace
\textbf{Floral} & Flower length (mm) & \textbf{Floral} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube}\\
\addlinespace
\textbf{Floral} & Inflorescence width (mm) & \textbf{Floral} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic}\\
\addlinespace
\textbf{Floral} & Style length (mm) & \textbf{Floral} & Nectar & \makecell[l]{Presence \\ Absence}\\
\addlinespace
\textbf{Floral} & Ovules per flower & \textbf{\makecell[c]{Reproductive \\ biology}} & Autonomous selfing & \makecell[l]{None \\ Low \\ Medium \\ High}\\
\addlinespace
\textbf{Floral} & Flowers per plant & \textbf{\makecell[c]{Reproductive \\ biology}} & Compatibility system & \makecell[l]{Self-incomp. \\ Part. self-comp. \\ Self-comp.}\\
\addlinespace
\textbf{Floral} & Nectar ($\mu$l) & \textbf{\makecell[c]{Reproductive \\ biology}} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious}\\
\addlinespace
\textbf{Floral} & Nectar (mg) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Nectar concentration ($\%$) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Pollen grains per flower & \textbf{} &  & \\
\addlinespace
\textbf{\makecell[c]{Reproductive \\ biology}} & Autonomous selfing (fruit set) & \textbf{} &  & \\
\bottomrule
\end{tabular}
\end{table}
